Cudele: A File System with Programmable Consistency and Durability

### 2017-03-28 paper/sections/4_methodology.tex
paper: add programmability section
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

% describe the figure
Figure~\ref{fig:decouple} shows the mechanisms (labeled arrows) in Cudele and
which entity they are performed by (gray boxes). The metadata store and journal
are different ways of representing the namespace. The metadata store represents
the namespace as a tree of directory fragments and is easier to read/traverse.
The journal represent the namespace as a list of events. It is
a ``pile system"; writes are fast but reads are slow because state must be
reconstructed.  Specifically, reads are slow because there is more data to
read, it is unorganized, and many of the updates may be redundant.

% describe the mechanisms
Cudele presents 6 mechanisms: RPCs, Stream, Create, Volatile Apply, Local Persist, and Global
Persist. ``RPCs" does round trip remote procedure calls to establish
consistency; it is the default implementation for complying with POSIX in
CephFS. ``Stream" has the metadata servers stream a journal of metadata updates
into the object store. ``Create" allows clients to append metadata events to an
in-memory journal. ``Volatile apply" takes the in-memory journal on the client
and applies it directly to the in-memory metadata store of the metadata server
cluster. ``Local Persist" takes the in-memory journal and writes it to the client's
disk. ``Global Persist" saves the journal as a an object in the object store from the
client.

Next, we discuss how these mechanisms can be composed to get different
consistency and durability semantics.

@@

% what is programmability
A programmable storage system exposes internal subsystem as building blocks for
higher level services. This `dirty-slate' approach limits redundant code and
leverages the robustness of the underlying storage system. Cudele uses this
approach and re-uses some of the building blocks from the Malacology
programmable storage system~\ref{sevilla:eurosys17} to great success and
requires only:

@@

% how it works in CephFS, how I used it
\subsubsection{Metadata Store}

\subsubsection{Journal}

\subsubsection{Journal Tool}
The journal tool is used for disaster recovery and lets administrators view and
modify the journal of metadata updates; it can read the journal, erase events
from the journal, and apply the updates in the journal to the metadata store.
To apply journal updates to the metadata store, the journal tool reads the
journal segments from object store objects and applies the update to the
metadata store (which are also stored as object store objects).

% Journal import
The journal tool imports journals from binary files stored on disk.  First the
header of the dump is sanity checked and written into RADOS to the ``header"
object.  The ``header" object has metadata about the journal as well as the
locations of all the journal pointers (e.g., where the tail of the journal is,
where we are currently trimming, etc.).  Then the journal events are cleaned
(erasing trailing events that are not part of the header) and written as
objects into RADOS.  Note that while the journal is in RADOS, the metadata
servers do do not have the namespace reconstructed in memory so the metadata
cluster will not service requests relating to the journal of imported events.
To construct the namespace in the collective memory of the metadata servers we
need to first construct the namespace in RADOS. The journal tool can explicitly
do this by  applying the journal to the metadata store in RAODS. This will pull
the objects containing journal segments and replay them on the metadata store.
Finally, we delete the journal in RADOS and restart the metadata servers so
they rebuild their caches.

% Journal export
The journal tool exports journals to binary files stored on disk. First the
journal is scanned for the header and then journal is recovered. To recover the
journal the ``header" object is read off disk and then objects are probed in
order and starting from the write position saved in the header. Probing will
update the write position if it finds objects with data in them.

% Journal export
When exporting a journal of events, the journal tool first scans the journal to
check for corruption. Then it recovers the journal by reading the ``header"
object out of RADOS.  After reading the header, the journal tool can pull
journal segments from RADOS because it knows how many objects to pull and how
far to seek within those objects.

@@

\subsubsection{Inode Cache}

\subsubsection{Large Inodes}

% file type
To assign consistency and durability to the subtrees we store policies in
the directory inode. This approach uses the File Type interface from the
Malacology programmable store system~\cite{sevilla:eurosys17} and it tells clients how to access
the underlying data or metadata. The underlying implementation stores
executable code in the inode that calls the different Cudele mechanisms. Of
course, there are many security and access control aspects of this approach but
that is beyond the scope of this paper.

>>>>

% describe the figure
Figure~\ref{fig:decouple} shows the mechanisms (labeled arrows) in Cudele and
which entity they are performed by (gray boxes). The metadata store and journal
are different ways of representing the namespace.  Cudele presents 6
mechanisms: RPCs, Stream, Create, Volatile Apply, Local Persist, and Global
Persist. ``RPCs" does round trip remote procedure calls to establish
consistency; it is the default implementation for complying with POSIX in
CephFS. ``Stream" has the metadata servers stream a journal of metadata updates
into the object store. ``Create" allows clients to append metadata events to an
in-memory journal. ``Volatile apply" takes the in-memory journal on the client
and applies it directly to the in-memory metadata store of the metadata server
cluster. ``Local Persist" takes the in-memory journal and writes it to the
client's disk. ``Global Persist" saves the journal as a an object in the object
store from the client.  Next, we discuss how these mechanisms can be composed
to get different consistency and durability semantics.

@@

% what is programmability
A programmable storage system exposes internal subsystem as building blocks for
higher level services. This `dirty-slate' approach limits redundant code and
leverages the robustness of the underlying storage system. Cudele uses this
approach and re-uses some of the building blocks from the Malacology
programmable storage system~\cite{sevilla:eurosys17} to great success and
requires only:

@@

% how it works in CephFS, how I used it
\subsubsection{Metadata Store}
\label{sec:metadata-store}

% how it works in CephFS
In CephFS, the metadata store is a data structure that represents the file
system namespace. This data structure is stored in two places: in memory ({\it
i.e.} in the collective memory of the metadata server cluster) and as objects
in the object store. In the object store, directories and their inodes are
stored together in objects to improve the performance of scans.  The metadata
store data structure is structured as a tree of directory fragments making it
easier to read and traverse.\\

% how i used it
\noindent\textbf{Cudele}: uses the metadata store format to write objects
formatted like the journal into the object store. By writing with the same
format, the metadata servers can read and use the recovery code to materialize
the updates from a client's decoupled namespace ({\it i.e.} merge).

\subsubsection{Journal}
\label{sec:journal}

The journal is the second way that CephFS represents the file system namespace;
it is a log of metadata updates that can materialize the namespace when the
updates are replayed onto the metadata store. The journal is a ``pile system";
writes are fast but reads are slow because state must be reconstructed.
Specifically, reads are slow because there is more state to read, it is
unorganized, and many of the updates may be redundant.\\

\noindent\textbf{Cudele}: uses the journal subsystems within the metadata
server to replay updates onto the in-memory metadata store.  When the clients
are ready to merge their updates back into the global namespace, they pass a
binary file of  metadata updates to the metadata server. The metadata server
uses its recovery code to ``replay" the updates.

\subsubsection{Journal Tool}
\label{sec:journal-tool}

The journal tool is used for disaster recovery and lets administrators view and
modify the journal. It can read the journal, export the journal as a file,
erase events, and apply updates to the metadata store.  To apply journal
updates to the metadata store, the journal tool reads the journal from object
store objects and replays the updates on the metadata store in the object
store.\\

\noindent\textbf{Cudele}: using the journal tool, clients can materialize the
journal in memory and append events to their own copy. When they are done they
export their journal to a binary file. Clients also use the journal tool to
save their updates to disk or to the object store.

% Journal import
%The journal tool imports journals from binary files stored on disk.  First the
%header of the dump is sanity checked and written into RADOS to the ``header"
%object.  The ``header" object has metadata about the journal as well as the
%locations of all the journal pointers (e.g., where the tail of the journal is,
%where we are currently trimming, etc.).  Then the journal events are cleaned
%(erasing trailing events that are not part of the header) and written as
%objects into RADOS.  Note that while the journal is in RADOS, the metadata
%servers do do not have the namespace reconstructed in memory so the metadata
%cluster will not service requests relating to the journal of imported events.
%To construct the namespace in the collective memory of the metadata servers we
%need to first construct the namespace in RADOS. The journal tool can explicitly
%do this by  applying the journal to the metadata store in RAODS. This will pull
%the objects containing journal segments and replay them on the metadata store.
%Finally, we delete the journal in RADOS and restart the metadata servers so
%they rebuild their caches.

% Journal export
%First the
%journal is scanned for the header and then journal is recovered. To recover the
%journal the ``header" object is read off disk and then objects are probed in
%order and starting from the write position saved in the header. Probing will
%update the write position if it finds objects with data in them.

% Journal export
%When exporting a journal of events, the journal tool first scans the journal to
%check for corruption. Then it recovers the journal by reading the ``header"
%object out of RADOS.  After reading the header, the journal tool can pull
%journal segments from RADOS because it knows how many objects to pull and how
%far to seek within those objects.

@@

\subsubsection{Inode Cache}
\label{sec:inode-cache}

In CephFS, the inode cache reduces the number of RPCs between clients and
metadata servers. Without contention clients can resolve metadata reads locally
thus reducing operations like \texttt{lookup()}. The size of the inode cache is
configurable so as not to saturate the memory on the metadata server -- inodes
in CephFS are about 1400
bytes\footnote{http://docs.ceph.com/docs/jewel/dev/mds\_internals/data-structures/}.
The inode cache has code for manipulating inode numbers, such as pre-allocating
inodes to clients.\\

\noindent\textbf{Cudele}: uses the internal inode cache code to allocate inodes
to clients that decouple parts of the namespace and to skip inodes used by the
client at merge time.

\subsubsection{Large Inodes}
\label{sec:large-inodes}

In CephFS, inodes store policies for policies, like how the file is striped
across the object store or for managing subtrees for load balancing. These
policies adhere to logical partitionings of metadata or data, like Ceph pools and file
system namespace subtrees. To implement this, the namespace data structure has
the ability to recursively apply policies to subtrees and to isolate subtrees
from each other.  \\

\noindent\textbf{Cudele}: uses the large inodes to store consistency and
durability policies in the directory inode. This approach uses the File Type
interface from the Malacology programmable store
system~\cite{sevilla:eurosys17} and it tells clients how to access the
underlying metadata. The underlying implementation stores executable code in
the inode that calls the different Cudele mechanisms. Of course, there are many
security and access control aspects of this approach but that is beyond the
scope of this paper.
